---
title: "Statistical Inference Course Project Analysis"
author: "Justin Meyer"
date: "December 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This project will analyze the ToothGrowth dataset. 

## Load the ToothGrowth data and perform some basic exploratory data analyses

What columns are in the dataset? What data types are those columns?

```{r}

tooth <- ToothGrowth

str(tooth)

```

How many rows are in the dataset?

```{r}

nrow(tooth)

```

## Provide a basic summary of the data

Below is a six number summary of the data:

```{r}

summary(tooth)

```

The three dose sizes are evenly distributed across the two supplements:
```{r}

table(tooth$supp, tooth$dose)

```

Length of growth and dose are strongly positively correlated:

```{r}

cor(tooth$len, tooth$dose)

```

The distribution of the length variable seems reasonable with length appearing to possibly be longer for supplement OJ.

```{r}

library(ggplot2)
ggplot(tooth, aes(len, colour = supp)) +
  geom_density() + 
  labs(title = "Distribution of Length")

```

Below is a chart showing the relationship between length and dose, broken out by supplement type. At the lower two doses (0.5 and 1.0) OJ seems to promote more tooth growth than VC but at the highest dose (2.0) the two supplements seem to produce similar growth.

```{r}

ggplot(data=tooth, aes(x=dose, y=len, group=supp, colour=supp)) +
    geom_point() + 
    labs(title = "Tooth Growth by Supplement Type and Dose")

```

## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

An independent samples t-test comparing the two supplements shows that while there is a difference between the supplements OJ and VC, the difference is not significant at the 0.05 level.
```{r, message=FALSE}

t.test(tooth$len ~ tooth$supp)

```

## State your conclusions and the assumptions needed for your conclusions.

Based on the above analysis I conclude that supplement type does not have a significant impact on tooth growth. The independent samples t-test that was employed requires the following assumptions:  
1. Each observation is independent.  
2. The length data follow a normal distribution.  
3. The variation of the length data in both populations is similar.  
