---
title: "Statistical Inference Course Project Analysis"
author: "Justin Meyer"
date: "December 21, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This project will analyze the toothGrowth dataset. 

## Load the ToothGrowth data and perform some basic exploratory data analyses

Simulate the exponential distribution 1000 times with 40 samples and a rate of 0.2. Take both the means and variances of those 1000 distributions.

```{r}

# Simulate means of the exponential distribution
means <- NULL
for(i in 1:1000) means <- c(means, mean(rexp(n = 40, rate = 0.2)))

# Simulate variances of the exponential distribution
vars <- NULL
for(i in 1:1000) vars <- c(vars, var(rexp(n = 40, rate = 0.2)))

```

## Sample Mean versus Theoretical Mean

*Did you show where the distribution is centered at and compare it to the theoretical center of the distribution?* 

As per the instructions "The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations."

Therefore the theoretical mean of the distribution is 1/0.2 = 5. The following chart compares the observed mean (blue) to the theoretical mean (red).

As you can see the observed mean `r mean(means)` is very close to the theoretical mean 5:

```{r, message=FALSE}

# Create a plot
means <- data.frame(means)

library(ggplot2)
ggplot(data = means, aes(means)) + 
        geom_histogram() + 
        geom_vline(aes(xintercept = (mean(means))),
            color="blue", linetype = "solid", size = 1) + 
        geom_vline(aes(xintercept = 5),
            color="red", linetype = "dashed", size = 1) + 
        labs(title = "Exponential Distribution Means vs. Theoretical Mean")

```

## Sample Variance versus Theoretical Variance

*Did you show how variable it is and compare it to the theoretical variance of the distribution?*

As per the instructions "The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations."

Therefore the theoretical standard deviation of the distribution is 1/0.2 = 5. Since the standard deviation is the square root of variance the theoretical variance must be 5^2 = 25. The following chart compares the observed variance (blue) to the theoretical variance (red).

The observed variance of the exponential distribution `r mean(vars)` is very close to the theoretical variance 25.

```{r, message=FALSE}

# Create a plot
vars <- data.frame(vars)

ggplot(data = vars, aes(vars)) + 
        geom_histogram() + 
        geom_vline(aes(xintercept = (mean(vars))),
            color="blue", linetype = "solid", size = 1) + 
        geom_vline(aes(xintercept = 25),
            color="red", linetype = "dashed", size = 1) + 
        labs(title = "Exponential Distribution Variances vs. Theoretical Variance")

```

## Distribution

Simulate the random exponential distribution.

```{r, message=FALSE}

# Create data
rexps <- rexp(n = 1000, rate = 0.2)

```

The following chart shows the random exponential distribution. As you can see it is not normal, with the distribution skewed to the right.

```{r, message=FALSE}

# Create a plot

rexps <- data.frame(rexps)

ggplot(data = rexps, aes(rexps)) + 
        geom_histogram() + 
        labs(title = "Random Exponential Distribution")

```

The means of 1000 random exponential distributions are normal though. 

```{r, message=FALSE}

# Create a plot
ggplot(data = means, aes(means)) + 
        geom_histogram() + 
        labs(title = "Distribution of 1000 Exponential Distribution Means")

```
